Formalization of Abstract Rewriting

### 2025-02-22 Relations/Paper/Formalization.tex
Formalization update

\section{Formalization of Terese}
\label{sec:Formalization}
The original ambition of this project was to formalize the contents of the \emph{abstract reduction 
systems} chapter in Term Rewriting Systems [TeReSe], a standard text in rewriting. 
The motivation for the project was to provide construvtive proofs in Agda 
of the propositions and theorems in the chapter. This in turn provides a foundation for 
future work in programming languages theory.

In order to formalize the chapter we have aimed to provide constructive counterparts to the 
classical proofs found in the text, and to make explicit any classical properties we depend 
on when a constructive proof could not be found. It is from this formalization that our 
original work has emerged. We have gone beyond \terese in fleshing out the interrelationship 
of the properties most interesting from the perspective of programming languages theory (namely 
being confluent and strongly normalizing). 
For some properties we suggest alternatives which are more conducive to a constructive project. 
We look into many notions of well-foundedness and again explore their interrelationship.

This section explains the deviations we took from \terese and lays the groundwork for the original work 
that we have carried out.

\sacomment{Maybe mention where we have stayed true to the book, but I don't think this is key}

Our first deviation from \terese is in defining strong normalization. The book defines 
an element as strongly normalizing if every reduction sequence starting from it is finite. 
We wanted to avoid introducing the notion of finiteness to our definition and so we 
followed the alternative definition given in \terese : a relation is strongly normalizing when its 
converse is well founded. In \terese a particular definition of well-foundedness is given but 
it is one with which we were not satisfied and so we explored different definitions of 
well-foundedness. This exploration is detailed in section \ref{sec:Well-foundedness}.

Another deviation we have taken from \terese relates to the use of the definition of an 
increasing relation (see dfn:\ref{d:inc}). This property was one we found particularly difficult 
to make use of in a constructive setting (for example showing that it along with the properties 
of being WCR and WN implies SN) and so we suggest an alternative property, $\RP$.
(dfn:\ref{d:RP}).

>>>>

\section{Formalization of Terese}
\label{sec:Formalization}
This sections acts as a guide of our formalization of \terese and explains the deviations we took.

\subsection{Formalization of Definitions}
We have formalized those definitions which were necessary for the formalization of the propositions and theorems found 
in Chapter 1 of \terese. We use Agda's built-in \texttt{Set} for sets and define a relation in \texttt{Predicates.agda}.
The closure operators are formalized in \texttt{ARS-Closure.agda}. 
The definitions given in \terese in \texttt{1.1.8 DEFINITION (confluence)} 
are formalized in \texttt{ARS-Base.agda}. We only cover $\WCR$ and $\CR$ of the definitions from this section 
as the other definitions were not involved in the more interesting formalizations.

The definitions given in \terese in \texttt{1.1.13 DEFINITION (normalization)} 
are formalized in \texttt{ARS-Properties.agda}.

The following is a summary of the differences in definition we have used as mentioned in Section \ref{sec:Definitions}. 
\begin{description}
    \item[$\SN$] We make use of the well-founded defintion, rather than the finite reduction sequence definition.
    \item[$\NPe$] We use the equivalent $\NP$ definition. 
    \item[$\Inc$] We use the more general property $\RP$.      
\end{description}

Our motivation for using $\RP$ rather than $\Inc$ is that we found $\Inc$ to be a particularly difficult property 
to use in a constructive setting.

\subsection{Propositions}
There are three sets of propositions given in \terese: \texttt{1.1.9}, \texttt{1.1.10}, \texttt{1.1.11}. 
The formalization of the proofs are in \texttt{ARS-Propositions.agda}. These formalizations were straightforward 
and as such are not the focus of this paper.

\subsection{Theorems}
\subsubsection{Newman's Lemma}
There are three proofs of Newman's Lemma ($\SN \land \WCR \implies \CR$) given in \terese. 
The first and third of these proofs make use of classical properties,
